---
title: "Bellabeat case study"
author: "Krishna pareek"
date: "2024-07-18"
output: html_document 
---
### install and load packages 
#### tidyverse = wrangle the data
#### skimr     = for summary
#### here      = manage file path   
#### janitor   =simlify data cleaning
#### ggplot2   = for visualization
#### lubridate = for date and time
#### dplyr     = for cleaning
#### sqldf     = sql function
#### plotrix   = 3d pie chart
```{r}
chooseCRANmirror(ind = 1)
```

```{r}
#setting the environment for .R
install.packages("tidyverse") 
install.packages("skimr")
install.packages("here")
install.packages("janitor")
install.packages("ggplot2") 
install.packages("lubridate")
install.packages("dplyr")             
install.packages("sqldf")
install.packages("plotrix")

library(tidyverse)
library(skimr)
library(here)
library(janitor)
library(ggplot2)  
library(lubridate)
library(dplyr)              
library(sqldf)
library(plotrix)
```

### get and set the data path for importing required files

```{r}
getwd()
setwd("C:/Users/Hp-D/Desktop/bellabeat case study/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16")
daily_activity <- read.csv("dailyActivity_merged.csv")
sleep_day <- read.csv("sleepDay_merged.csv")
weight <- read.csv("weightLogInfo_merged.csv")
hourly_step <- read.csv("hourlySteps_merged.csv")
```

#### take a closer look to the data clean it remove the duplicate and
#### NA value.
```{r}
head(daily_activity)
head(sleep_day)
head(weight)
head(hourly_step)
```


```{r}
sum(is.na(daily_activity))
sum(is.na(sleep_day))
sum(is.na(weight))

sum(duplicated(daily_activity))
sum(duplicated(sleep_day))
sum(duplicated(weight))
sum(duplicated(hourly_step))
```
####here NA is "fat" at different date so we leave it. 
#### and remove the duplicate entries


```{r}
weight_info <- weight %>% select( Id , Date , WeightKg , WeightPounds , BMI , IsManualReport, LogId)
head(weight_info)

sleep_day <- sleep_day[!duplicated(sleep_day), ]
sum(duplicated(sleep_day))
```

### seprate the date and time 

```{r}
sleep_day_new <- sleep_day %>% separate(SleepDay, c("date", "time"), " ")
head(sleep_day_new)

weight_info_new <- weight_info %>% separate(Date, c("date", "time"), " ")
head(weight_info_new)

```

### find the unique row

```{r}
n_distinct(daily_activity$Id)
n_distinct(sleep_day_new$Id)
n_distinct(weight_info_new$Id)
n_distinct(hourly_step$Id)
```
#### by this we know that here is 33 unique id in Daily_activity , 24 in sleep_day_new & 8 in weight_info_new

### change the data formate.
```{r}
daily_activity$ActivityDate = as.Date(daily_activity$ActivityDate, "%m/%d/%Y") 
glimpse(daily_activity)

sleep_day_new$date = as.Date(sleep_day_new$date, "%m/%d/%y")
glimpse(sleep_day_new)

weight_info_new$date = as.Date(weight_info_new$date, "%m/%d/%y")
glimpse(weight_info_new)
```

###find out the centrel trend of entire dataset of
#### daily_activity, sleep_day_new, weight_info_new
```{r}
daily_activity %>% 
  select(TotalSteps, TotalDistance, VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes,
         Calories) %>% summary()
  
sleep_day_new %>% select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>% summary()

weight_info_new %>% select(WeightKg, WeightPounds, BMI) %>% summary()
```

### add a new column
#### for weekday

```{r}
daily_activity <- daily_activity %>% mutate( Weekday = weekdays(as.Date(ActivityDate, "%m/%d/%Y")))
```

### merge data
```{r}
merge1 <- merge(sleep_day_new, weight_info_new, by = "Id", all = TRUE )
merge_data <- merge(merge1, daily_activity, by = "Id", all = TRUE)
```
```{r}
str(merge_data)
```

### set order
```{r}
merge_data$Weekday <- factor(merge_data$Weekday, levels =  c("Monday", 
                       "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

ordered_data <- merge_data[order(merge_data$Weekday), ]

head(ordered_data)
```

### check for NA and duplicate in merge data
```{r}
sum(is.na(merge_data))
sum(duplicated(merge_data))
n_distinct(merge_data$Id)
```

### creat visual for total active minute and weekday
```{r}
daily_activity$total_active_minitus <- daily_activity$VeryActiveMinutes+daily_activity$FairlyActiveMinutes+daily_activity$LightlyActiveMinutes


ggplot(daily_activity, aes(x= Weekday, y = total_active_minitus)) +
  geom_bar(stat = "identity", fill = "green") +
  labs(x = "Weekday", y = "Total Active Minutes", title = "Total Active Minutes per Weekday")

```

###as we see people are mostly active on the mid day of the week

### Time Spent Awake vs. Total Sleep Duration in Bed
```{r}
sleep_day_new$awake_time_in_bed <- sleep_day_new$TotalTimeInBed - sleep_day_new$TotalMinutesAsleep

ggplot(sleep_day_new, aes(x=TotalMinutesAsleep, y= awake_time_in_bed))+
  geom_point(color = "blue") +
  geom_smooth(color = "orange")+
  labs(x = "time as sleep", y = "awaken time in bed", title = "relation between sleep and awaken time in bed")

```

### hourly steps
```{r}
head(hourly_step)
n_distinct(hourly_step$Id) #33 users

hourly_step$ActivityHour=as.POSIXct(hourly_step$ActivityHour,format="%m/%d/%Y %I:%M:%S %p")
hourly_step$Hour <-  format(hourly_step$ActivityHour,format= "%H")
head(hourly_step)

ggplot(data=hourly_step, aes(x=Hour, y=StepTotal, fill=Hour))+
  geom_bar(stat="identity")+
  labs(title="Hourly Steps")
```

### from  5 to 7 the people or the users are more active


####take the sum of (sedentary, lightly, fairly, active ) minitus, and make a variable name -> active_minitus
```{r}
Sedentary <- sum(daily_activity$SedentaryMinutes)
Lightly <- sum(daily_activity $LightlyActiveMinutes)
Fairly <- sum(daily_activity $FairlyActiveMinutes)
Active <- sum(daily_activity $VeryActiveMinutes)

activity_minutes <- c(Sedentary,Lightly, Fairly, Active)
```

#### calculate the percentage of active_minitus , for each level of activity
```{r}
activity_percent <- round(activity_minutes/sum(activity_minutes)*100,1)
```

###create a pie chart showing the percentage of active minutes for each level of activity
```{r}
par(bg= "white")
legend_labels <- c("Sedentary","Lightly Active", "Fairly Active","Very Active")
pie3D(activity_percent, labels=paste0(activity_percent,"%"),
      main="Percentage of Active Minutes by Activity Level", 
      col=c("darkred","purple", "yellow","green"),
      border="lightgrey", labelcex = 0.7)
legend("topright", legend_labels, cex=0.9,
       fill=c("darkred","purple", "yellow","green"),x.intersp = 1, y.intersp = 1)
```


